---
title: "Deployment"
author: "Javier López, Borja García, Alexandru Neamtu"
date: "2024-05-23"
output: html_document
---

# Exercise 5.- Model-agnostic: Partial Dependency Plot (PDP).

## 1.- One dimensional Partial Dependence Plot.
The partial dependence plot shows the marginal effect of a feature on the predicted outcome of a previously fit model.
### EXERCISE:
Apply PDP to the regression example of predicting bike rentals. Fit a random forest approximation for the prediction of bike rentals (*cnt*). Use the partial dependence plot to visualize the relationships the model learned. Use the slides shown in class as model.2


```{r}
library(dplyr)
library(plotly)
library(reshape2)
library(lubridate)
library(randomForestSRC)

days <- read.csv("day.csv")
hour <- read.csv("hour.csv")
days$dteday <- as_date(days$dteday)
days_since <- select(days, workingday, holiday, temp, hum, windspeed, cnt)
days_since$days_since_2011 <- int_length(interval(ymd("2011-01-01"), days$dteday)) / (3600*24)
days_since$SUMMER <- ifelse(days$season == 3, 1, 0)
days_since$FALL <- ifelse(days$season == 4, 1, 0)
days_since$WINTER <- ifelse(days$season == 1, 1, 0)
days_since$MISTY <- ifelse(days$weathersit == 2, 1, 0)
days_since$RAIN <- ifelse(days$weathersit == 3 | days$weathersit == 4, 1, 0)
days_since$temp <- days_since$temp * 47 - 8
days_since$hum <- days_since$hum * 100
days_since$windspeed <- days_since$windspeed * 67
rf <- rfsrc(cnt~., data=days_since)
results <- select(days_since, days_since_2011, temp, hum, windspeed, cnt)


nr <- nrow(days_since)
for(c in names(results)[1:4])
{
  for(i in 1:nr){
    r <- days_since
    r[[c]] <- days_since[[c]][i]
    sal <- predict(rf, r)$predicted
    results[[c]][i] <- sum(sal) / nr
  }
}
p1 = ggplot(days_since, aes(x = temp, y = results$temp)) + ylim(0, 6000) + geom_line() + geom_rug(sides="b", alpha = 0.5) + labs(x = "Temperature")
p2 = ggplot(days_since, aes(x = hum, y = results$hum)) + ylim(0, 6000) + geom_line() + geom_rug(sides="b", alpha = 0.5) + labs(x = "Humidity")
p3 = ggplot(days_since, aes(x = days_since_2011, y = results$days_since_2011)) + ylim(0, 6000) + geom_line() + geom_rug(sides="b", alpha = 0.5) + labs(x = "Days since 2011", y = "Predictions")
p4 = ggplot(days_since, aes(x = windspeed, y = results$windspeed)) + ylim(0, 6000) + geom_line() + geom_rug(sides="b", alpha = 0.5) + labs(x = "Windspeed")
#p1
subplot(p3, p1, p2, p4, titleX = TRUE, titleY = TRUE, shareX = FALSE, shareY = TRUE)
```

### QUESTION:
Analyse the influence of *days since 2011, temperature, humidity and wind speed* on the predicted bike counts.

Days since 2011: as time goes by, there is generally a clear upward trend in bicycle rentals, although in the last few days the predictions are decreasing. 

Temperature: The number of rented bicycles increases with the temperature up to around 20 degrees Celsius, but then decreases as the temperature rises from around 25 degrees Celsius onwards. 

Humidity: up to 50% humidity, the number of rented bicycles remains constant, but then decreases inversely to the increase in humidity. 

Windspeed: the prediction for bicycle rentals decreases gradually until a wind speed of approximately 23 km/h is reached. At this point, the predictions are constant.


## 2.- Bidimensional Partial Dependency Plot.
### EXERCISE:
Generate a 2D Partial Dependency Plot with humidity and temperature to predict the number of bikes rented depending on those parameters.

*BE CAREFUL*: due to the size, extract a set of random samples from the BBDD before generating the data for the Partial Dependency Plot.
Show the density distribution of both input features with the 2D plot as shown in the class slides.

*TIP*: Use geom_tile() to generate the 2D plot. Set width and height to avoid holes.

```{r}
sampled <- sample_n(days_since, 40)
temp <- sampled$temp
hum <- sampled$hum
th <- inner_join(data.frame(temp),data.frame(hum), by=character())
th$p <- 0
for(i in 1:nrow(th)){
  r <- days_since
  r[["temp"]] <- th[["temp"]][i]
  r[["hum"]] <- th[["hum"]][i]
  
  sal <- predict(rf, r)$predicted
  th[["p"]][i] <- sum(sal) / nr
}
#View(th)
```

```{r}
p5 = ggplot(th, aes(x = temp, y = hum, fill = p)) + geom_tile(width = 10, height = 15) + geom_rug(alpha = 0.5) + labs(x = "Temperature", y = "Humidity", fill = "Number of bikes")
p5
```

### QUESTION:
Interpret the results.

In this graph we can see that the highest number of bicycles are rented when the temperature is between 15 and 20 degrees and the relative humidity is between 0 and 70%, while the lowest number of bicycles are rented when the humidity is very high and the temperature is very low. 

## 3.- PDP to explain the price of a house.
### EXERCISE:
Apply the previous concepts to predict the *price* of a house from the database *kc_house_data.csv*. In this case, use again a random forest approximation for the prediction based on the features *bedrooms, bathrooms, sqft_living, sqft_lot, floors and yr_built.*
Use the partial dependence plot to visualize the relationships the model learned.
*BE CAREFUL*: due to the size, extract a set of random samples from the BBDD before generating the data for the Partial Dependency Plot.

```{r}
house_data<- read.csv("kc_house_data.csv")
sampled <- sample_n(house_data, 1000)
sampled <- select(sampled, bedrooms, bathrooms, sqft_living, sqft_lot, floors, yr_built, price)
rfsrc <- rfsrc(price~., data=sampled)
res <- select(sampled, bedrooms, bathrooms, sqft_living, floors, price)
rows <- nrow(sampled)
for(name in names(res)[1:4])
{
  for(i in 1:rows){
    resul <- sampled
    resul[[name]] <- sampled[[name]][i]
    pred <- predict(rfsrc, resul)$predicted
    res[[name]][i] <- sum(pred) / rows
  }
}
```

```{r}
figure6 = ggplot(sampled, aes(x = bedrooms,y = res$bedrooms)) + geom_line() + geom_rug(sides = "b", alpha = 0.5) + labs(x = "Bedrooms",y = "Prediction")
figure7 = ggplot(sampled,aes(x=bathrooms,y=res$bathrooms)) + geom_line() +  geom_rug(sides = "b", alpha = 0.5) + labs(x = "Bathrooms", y = "")
figure8 = ggplot(sampled,aes(x=sqft_living,y=res$sqft_living)) + geom_line() +  geom_rug(sides = "b", alpha = 0.5) + labs(x = "Sqft Living", y = "")
figure9 = ggplot(sampled,aes(x=floors,y=res$floors)) +   geom_line() + geom_rug(sides = "b", alpha = 0.5) + labs(x = "Floors", y = "")
subplot(figure6, figure7, figure8, figure9, titleX = TRUE, titleY = TRUE, shareX = FALSE, shareY = FALSE)
```


### QUESTION:
Analyse the influence of *bedrooms, bathrooms, sqft_living and floors* on the predicted price.


As for the bedrooms, the price of the house has a positive correlation with the number of bedrooms in the house, i.e., as the number of bedrooms in the house increases, the price of the house also increases.

In the other hand, the bathrooms, the price of the house has a positive correlation with the number of bathrooms in the house, i.e., as the number of bathrooms in the house increases, the price of the house also increases.

Livable Square Footage, the price of the home has a positive correlation with the livable square footage of the home, i.e., as the livable square footage of the home increases, the price of the home also increases.

And finally, the floor, the price of the house has a positive correlation with the number of floors in the house, i.e., as the number of floors in the house increases, the price of the house also increases.